### 2024-11-23 paper/paper.qmd
updated paper.qmd

---
title: "Forecasting the Toronto Shelter System Performance on Reducing Chronic Homelessness "
subtitle: "My subtitle if needed"
author: 
  - Sarah Ding
thanks: "Code and data are available at: [https://github.com/sarahdingg/Toronto_Shelter_System_Forecast)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

# Load cleaned data
analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview
The dataset used for the current analysis is obtained from Open Data Toronto [@opendatatoronto], it is last refreshed on November 15, 2024. Open Data Toronto is an open source accessible to the public,

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Variables {#sec-variables}
The collected data from Open Data Toronto [@opendatatoronto] contains several key variables relevant to the analysis:

- **'date'**: The time period (month/year) for which the data is being published (format: mmm-yy)
- **'population_group'**: Represents different groups: all population, chronic, refugees, families, youth, single adult and non-refugees
- **'returned_to_shelter'**: People who were previously using the shelter system, then did not use the system for 3 months or longer, and have now returned. Some other communities may call this indicator “Returned from Inactive”
- **'newly_identified'**: People who entered the shelter system for the first time
- **'moved_to_housing'**: People who were using the shelter system and have moved to permanent housing.
- **'actively_homeless'**: People who have used the shelter system at least one time in the past three months and did not move to permanent housing.
- **'age'**: People's age recorded in the Shelter Management Information System, with ranges of under 16, 16-24, 25-34, 35-44, 45-54, 55-64, and over 65. 
- **'population_group_percentage'**: Represents the percentage of each population group (chronic, refugees, families, youth, single adult and non-refugees). The proportion is calculated from all the population.

For the purpose of the current analysis, only the chronic population group is included out of all population groups - for more details on the data cleaning process, refer to [Appendix - @sec-data-cleaning]

## Outcome variables

As the focus is to analyze and predict the performance of Toronto's shelter system in reducing homelessness, the primary outcome variable of interest is the number of individuals who have successfully transitioned to permanent housing, denoted as **‘moved_to_housing’**. The **`moved_to_housing`** variable captures the count of people who exited the shelter system and moved into stable, permanent housing arrangements. This variable serves as a direct measure of the shelter system's success in achieving its goal of reducing homelessness to rare, brief, and non-recurring instances. By examining and modeling the **`moved_to_housing`** variable, we can explore effectiveness of the shelter system and identify trends and predictors that influence successful housing transitions.

```{r}
#| label: fig-outcome-var
#| fig-cap: Bills of penguins
#| echo: false
#| 
# Add a year column to the dataset
analysis_data <- analysis_data %>%
  mutate(year = format(date, "%Y"))

# Group by year and calculate the average of moved_to_housing
summary_table <- analysis_data %>%
  group_by(year) %>%
  summarise(average_moved_to_housing = mean(moved_to_housing, na.rm = TRUE))

# Display the summary table
knitr::kable(summary_table, caption = "Average Moved to Housing by Year")
```

@fig-outcome-var presents

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## What did we learn about the world? talk about differences before pandemic and after

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix 1 {#sec-appendix1}

# Additional data details {#sec-additional-data-details}
how data is cleaned for analysis data and model data, why lagged variables are constructed for time-series analysis and regression models

## Data Cleaning {#sec-data-cleaning}

## Limitations and Improvements

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

# Apendix 2 {#sec-appendix2}

\newpage

# References